Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-11-01 Manuscript/manuscript.tex
Figure update2: done what was left to do in previou commit

The probabilities for a bound \ce{Ca^{2+}} ion to interact with
different oxygen moieties of POPC 
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O}$,
were analyzed by counting the contacts within the distance of $0.3 \, \mathrm{nm}$, 
as done previously in Ref.~\citenum{javanainen17}. 
%
%SAMULI: I commented out the average number of contacts. All the essential information
%for this is in the probabilities.
%
%In contrast to sodium, the dominant contribution
%to the binding of \ce{Ca^{2+}} to POPC membranes comes from the phosphate group. 
%% Numbers from simulation with 346mM conc. CaCl2
%The average number of contacts per lipid 
%between \ce{Ca^{2+}} at $290\,\mathrm{mM}$ concentration and any oxygen atom in POPC is 
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.134 $,    % = 17.1/128 
%whereas if only phosphate oxygen atoms are considered 
%the average number of contacts decreases only by a tiny amount to 
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.133 $,    % = 17.0/128 . 
%and if only carbonyl oxygen atoms are considered 
%%(phosphate oxygens contribute, but are not counted as contacts),
%this value is merely
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.033 $.    % = 4.3/128 . 
%
% Numbers from simulation with 980/845mM conc. CaCl2
%The average number of contacts per lipid between \ce{Ca^{2+}} and any oxygen atom in POPC is 
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 $,    % = 33.8/128 
%whereas if only phosphate oxygen atoms are considered 
%the average number of contacts decreases only by a tiny amount to 
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.262 $,    % = 33.5/128 . 
%and if only carbonyl oxygen atoms are considered 
%(phosphate oxygens contribute, but are not counted as contacts),
%this value is merely
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.066 $.    % = 8.4/128 . 
The results in Table \ref{tab:binding} show that almost all (99\%) of the 
bound Ca$^{2+}$ ions are in contact with the phosphate oxygens and a quarter
of them (25\%) interacts with the carbonyl oxygens. Thus, the interactions between
Calcium ions with only carbonyl oxygens are very unlikely (1\%), but significant
amount of the ions interact with both, phosphate and carbonyl oxygens.
The individual lipids or acyl chains were not identified in the analysis,
thus the simultaneous ion contacts with both oxygen moieties may be inter or intra molecular
and occur with the carbonyls in the either of the acyl chains.
The most likely contacts between Ca$^{2+}$ ions and phosphate oxygens is visualized with
the probability isodensity contours in Fig.~\ref{fig:volmaps}.
While the higher concentrations of \ce{CaCl2} naturally increased the amount of contacts per lipid,
%for $850\,\mathrm{mM}$ concentration of \ce{CaCl2}
%the average number of contacts per lipid is increased to
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 $, 
the distribution of contacts between phosphate and carbonyl oxygens was not affected.
%and density profiles of ions and water along the membrane normal in Fig.~\ref{fig:cacl-dens}.

>>>>

The probabilities for a bound \ce{Ca^{2+}} ion to interact with
different oxygen moieties of POPC 
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O}$,
were analyzed by counting the contacts within the distance of $0.3 \, \mathrm{nm}$, 
as done previously in Ref.~\citenum{javanainen17}. 
%
%SAMULI: I commented out the average number of contacts. All the essential information
%for this is in the probabilities.
%
%In contrast to sodium, the dominant contribution
%to the binding of \ce{Ca^{2+}} to POPC membranes comes from the phosphate group. 
%% Numbers from simulation with 346mM conc. CaCl2 (OPC3)
%The average number of contacts per lipid 
%between \ce{Ca^{2+}} at $290\,\mathrm{mM}$ concentration and any oxygen atom in POPC is 
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.134 $,    % = 17.1/128 
%whereas if only phosphate oxygen atoms are considered 
%the average number of contacts decreases only by a tiny amount to 
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.133 $,    % = 17.0/128 . 
%and if only carbonyl oxygen atoms are considered 
%%(phosphate oxygens contribute, but are not counted as contacts),
%this value is merely
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.033 $.    % = 4.3/128 . 
%
% Numbers from simulation with 980/845mM conc. CaCl2 (OPC3)
%The average number of contacts per lipid between \ce{Ca^{2+}} and any oxygen atom in POPC is 
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 $,    % = 33.8/128 
%whereas if only phosphate oxygen atoms are considered 
%the average number of contacts decreases only by a tiny amount to 
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.262 $,    % = 33.5/128 . 
%and if only carbonyl oxygen atoms are considered 
%(phosphate oxygens contribute, but are not counted as contacts),
%this value is merely
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.066 $.    % = 8.4/128 . 
The results in Table \ref{tab:binding} show that almost all (99\%) of the 
bound Ca$^{2+}$ ions are in contact with the phosphate oxygens,
with two thirds (67\%) are interacting with only phosphate oxygens
and the rest, one third (32\%), interacts also with the carbonyl oxygens. 
Thus, the interactions between
Calcium ions with only carbonyl oxygens are very unlikely (1\%), but significant
amount of the ions interact with both, phosphate and carbonyl oxygens.
The individual lipids or acyl chains were not identified in the analysis,
thus the simultaneous ion contacts with both oxygen moieties may be inter or intra molecular
and occur with the carbonyls in the either of the acyl chains.
The most likely contacts between Ca$^{2+}$ ions and phosphate oxygens is visualized with
the probability isodensity contours in Fig.~\ref{fig:volmaps}.
While the higher concentrations of \ce{CaCl2} naturally increased the amount of contacts per lipid,
%for $850\,\mathrm{mM}$ concentration of \ce{CaCl2}
%the average number of contacts per lipid is increased to
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 $, 
the distribution of contacts between phosphate and carbonyl oxygens was not affected.
%and density profiles of ions and water along the membrane normal in Fig.~\ref{fig:cacl-dens}.
